Notes on conformal manifolds

### 2025-06-13 Draft_MLFlatDirections.tex
update the draft

% ---- GRAPHIQUE ----
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[font=small,labelfont=bf,labelsep=space]{caption}
\usepackage{subfigure}
\captionsetup{width=0.9\textwidth,font={small,stretch=1.1}}
\addto\captionsenglish{\renewcommand{\figurename}{Fig.}}
\addto\captionsenglish{\renewcommand{\tablename}{Tab.}}
\definecolor{JoliBleu}{rgb}{0,0.55,0.55}
\definecolor{JoliVert}{rgb}{0.15,0.6,0}
\definecolor{JoliRouge}{rgb}{0.86,0.08,0}
\definecolor{JoliJaune}{rgb}{1,0.75,0}
\definecolor{JoliGris}{rgb}{0.52,0.52,0.51}
\definecolor{myblue}{RGB}{26, 77, 116}
\definecolor{myorange}{RGB}{181, 116, 30}
\definecolor{mydarkorange}{RGB}{166, 88, 0}
\definecolor{mygreen}{RGB}{21, 124, 80}
\definecolor{myblack}{RGB}{43, 65, 82}
\definecolor{myred}{rgb}{0.5, 0.0, 0.13}

@@

   \begin{Large}

\mbox{ \bfseries \boldmath  Notes: Machine learning flat directions}
   \end{Large}

@@

The AdS/CFT correspondence stands as one of the most profound dualities in theoretical physics, establishing a remarkable equivalence between gravitational theories in Anti-de Sitter (AdS) spacetimes and conformal field theories (CFTs) living on their boundaries. This correspondence has revolutionized our understanding of both quantum gravity and strongly coupled field theories, providing unprecedented insights into the holographic nature of gravity.

Within this holographic framework, supergravity theories in AdS backgrounds serve as the low-energy effective descriptions of string theory compactifications, making them natural laboratories for exploring the gravitational side of the duality. On the CFT side, the dual conformal field theories often exhibit rich structures of exactly marginal operators whose coupling constants parametrize conformal manifolds—moduli spaces where the theory remains conformal but with varying coupling strengths. The AdS/CFT correspondence maps these conformal manifolds to flat directions in the supergravity scalar potential, where the cosmological constant remains fixed while scalar field configurations vary continuously.

Understanding these flat directions is therefore of great importance for several interconnected reasons. First, they encode the conformal manifolds of the dual CFTs, providing direct access to the space of exactly marginal deformations and their associated beta functions. Second, they reveal the vacuum structure of supergravity theories, determining which background configurations preserve supersymmetry and their associated stability properties.

However, the explicit characterization of these flat directions presents formidable technical challenges. Supergravity scalar potentials, even in truncated models, typically involve dozens of scalar fields with intricate non-linear interactions. The resulting expressions for critical points—where all first derivatives vanish—quickly become too complex for traditional symbolic manipulation, rendering analytical approaches computationally intractable. This complexity paradox is particularly frustrating given that physically meaningful solutions often admit relatively simple parametric descriptions once discovered.

The emergence of machine learning techniques in theoretical physics offers new avenues for tackling such problems. Rather than attempting to solve the full symbolic system directly, we can leverage numerical methods to sample the solution space and then employ symbolic regression to extract analytical patterns from the resulting data. This hybrid approach potentially circumvents the computational bottlenecks that plague purely symbolic methods while maintaining the goal of obtaining exact analytical expressions.

In this work, we demonstrate the viability of this machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of three-dimensional chiral $\mathcal{N} = (1,1)$ supergravity arising from $AdS_3 \times S^3$ compactification from 6d supergravity, or from the compactification of type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the flat direction manifold with a novel annealed importance sampling algorithm for polynomial symbolic regression. We show that this approach successfully identifies the underlying three-dimensional conformal manifold and automatically discovers the polynomial constraints that characterize it analytically.

@@

As an initial simplification, we restrict our analysis to the scalar fields $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$, and $x_{10}$. This choice is made to facilitate a more tractable starting point for the investigation, and corresponds to a randomly selected subset of the original set of thirteen scalar fields. It is important to note that this subset no longer defines a consistent truncation of the full theory. Nevertheless, despite the fact that the effective theory defined by these five scalars may be formally ill-defined—due to potential higher-order interactions that could source the remaining scalar fields—the solutions we obtain remain valid solutions of the complete theory. This is because we compute the critical points of the scalar potential by first evaluating $\nabla V = 0$, and only then setting the remaining scalar fields to zero. By adopting this procedure—taking the gradient of the potential prior to truncation—we ensure that the resulting configurations satisfy the equations of motion and are thus genuine solutions of the full theory. Under this truncation, the scalar potential~\eqref{eq:scalarpotential} reduces to:
\begin{equation} \label{eq:scalarpotential124810}
	\begin{aligned}
		V &= \frac{1}{8}\, e^{-2\,\tilde{x}_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,\tilde{x}_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
		  & \quad\qquad\qquad - 4\,e^{3\,\tilde{x}_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad - 8\,e^{\tilde{x}_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
        & \quad\qquad\qquad + 4\,e^{2\,\tilde{x}_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,\tilde{x}_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
	\end{aligned}
\end{equation}

Let us make a comment at this point. Looking at the potential given in Eq.~\ref{eq:scalarpotential124810} and since we have an explicit formula, one might argue that the problem of identifying flat directions could be addressed by directly computing the gradient of the expression and subsequently simplifying the resulting equations. However, attempting to carry out this procedure using a symbolic solver such as Mathematica quickly reveals its limitations: the resulting expressions are too complex to be simplified into a manageable form, and do not yield usable relationships that express some variables in terms of others. This complexity does not, however, rule out the possibility that simpler solutions satisfying the condition of vanishing gradient may exist. In this work, we aim to identify such solutions, which we seek using numerical methods.

An additional advantage of this approach is its broader applicability. In scenarios where an explicit analytic expression for the potential is inaccessible—yet numerical evaluation remains feasible—our method retains its validity. Hence, the numerical strategy adopted here may prove effective even in cases where an exact analytic expression for the potential is out of reach.

@@

		To do so, we perform a basic gradient descent. We begin by randomly and uniformly initializing points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range $[-1,1]$. For this example, we choose to generate $10^5$ points. This number is motivated by the dimensionality of our problem. If we aim to populate all possible directions and want approximately $\mathcal{O}(10)$ points per direction, then we require $\mathcal{O}(10^5)$ points. Given that the true intrinsic dimensionality of the manifold is less than or equal to 5, this value serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

		We then perform gradient descent on the points using TensorFlow’s automatic differentiation. The loss function is defined as:
		\begin{equation}
			\mathcal{L} = \sum_{i = 1}^{n_{points}} || \nabla V(\vec{x}^{,i})||^2
		\end{equation}
		where $\vec{x}^{\,i} = (x^i_1,x^i_2,x^i_4,\tilde{x}^i_8,x^i_{10})$ denotes the data points.

		Through trial and error, we observed that periodically reinitializing the optimizer significantly improved the convergence rate. This technique bears similarity to the concept of warm restarts introduced in \cite{loshchilov2017sgdrstochasticgradientdescent}. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence.

		We interpret this as follows: when the optimizer is reinitialized, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.

		The evolution of the loss function, along with the learning rate schedule, is shown in Fig.~\ref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is reinitialized. We also observe that, following the last few reinitializations, the loss exhibits a small bump immediately after the restart.

		We interpret this behavior as a consequence of the increased learning rate: some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence.

		Although this strategy introduces the risk of desampling certain regions in favor of others, we did not observe any such issue in the remainder of our analysis.

@@

		The Adam optimizer was employed throughout the gradient descent procedure. It was reinitialized at iterations {250, 500, 750}, with the learning rate fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialized once more, this time with a reduced learning rate of $10^{-3}$. A final reinitialization was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 1000 epochs.

		Upon completion, it was observed that all data points converged to values close to $-4$. Specifically, nearly all data points attained values within an absolute error of $10^{-4}$, with the exception of four points whose deviations were on the order of $10^{-3}$. Similarly, 99,739 out of 100,000 data points exhibited gradient norms smaller than $10^{-3}$. These results are summarized in Figure~\ref{fig:Analyze_V_points}.

		\begin{figure}
			\centering
			\includegraphics[scale = 0.5]{Figures/Loss_grad_des.png}
			\caption{Evolution of the loss function during gradient descent and corresponding learning rate schedule, noted $\alpha$ on the left y-axis, both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
			\label{fig:loss_grad_des}
		\end{figure}

		As a first visualization, we present a triangular plot of the data in Fig. \ref{triangular_plot_1_2_4_8_10}. This figure shows all possible 2D projections of the data, along with the 1D histograms of each coordinate after gradient descent.

		\begin{figure}[t]
			\centering
			\includegraphics[scale = 0.3]{Figures/triangular_plot_1_2_4_8_10.png} %names need to be changed.
			\caption{Triangular plot showing $2d$ projections and $1d$ histograms of the data after the gradient descent}
			\label{triangular_plot_1_2_4_8_10}
		\end{figure}

		We can make a few comments on those plots. First, we observe non-trivial correlations in the data. Some of these may result from larger basins of attraction, such as in the $x_1/x_2$ plot. However, the structures seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots likely reflect genuine features of the manifold, and will show latter that this is indeed the case. We also note that all directions appear to be well populated. 
		%Furthermore, we can notice that a large majority of the points converge to the origin $\vec{0}$. This is not surprising, as this just mean that the round sphere solution has a larger basin of attraction, which sounds reasonable. We do not however attempt to shift the points along flat directions to depopulate the origin or repopulate other regions.

		%\paragraph{Local PCA: graphs of number of patches with given local dimension, and with varying size of local patch, discuss optimal patch size (we want at least few points in each direction)}

		Once the gradient descent has been completed and the flat direction sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we can first perform some exploratory analyses to better understand the data. Specifically, we aim to determine the dimensionality of the manifold and wether it consists of a single connected component or multiple disjoint components (e.g., two intersecting hyperplanes).

		To this end, we apply a local Principal Component Analysis (PCA). For each point, we identify its $k$ nearest neighbours and perform a PCA on that local neighbourhood. This procedure allows us to determine how many principal directions are needed to explain a given proportion $\epsilon$ of the data variance. In other words, it provides an estimate of the local dimensionality around each point.

		We perform this analysis for several values of $k$, namely $k \in \{5, 10, 20, 50, 100\}$, and we fix $\epsilon = 0.99$. The results are presented in Fig. \ref{local_pca}.

@@

		We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned a dimensionality of 2. This can be attributed to the fact that if the true dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality for a fraction of the points.

		Additionally, for $k \geq 20$, we observe an increasing number of points being assigned dimensionalities of 4 or 5. This behavior can be explained by the loss of locality when the number of neighbours becomes too large: increasing $k$ results in a coarser approximation, and the algorithm may then incorporate points that are no longer truly local. This artificial enlargement of the neighbourhood can cause the estimated local dimensionality to rise.

		We thus conclude that the manifold under investigation has an intrinsic dimension of 3.

@@

		Now we have sampled the manifold and extracted some basic information about it (namely its dimension and the fact that it is made of one block), we would like to see if we can extract some analytic formula to characterise it. What we have at the moment, are points on a 3-dimensional manifold, which are embedded in a 5-dimensional space. Therefore we can conclude that, in order to characterise the manifold, we need two independent constraints on the embedding coordinates $\vec{x} = (x_1,x_2,x_4,\tilde{x}_8,x_{10})$. If we have a look at the form of the potential \eqref{eq:scalarpotential124810}, we see that if we use $x_{8} = e^{\tilde{x}_8}$, that, up to a global $e^{-2\tilde{x}_8}$ factor, this potential is actually a polynomial of the embedding coordinates. Therefore, the components of $|\nabla V|$ are also polynomials in those variables. We conclude that the constraints on the embedding coordinates we are looking for are polynomial constraints of the form $p(\vec{x}) = 0$, and that there should be at least two of those. Note that we are now taking $\vec{x} = (x_1,x_2,x_4,\tilde{x}_8,x_{10})$. Of course if one takes directly the gradient of \eqref{eq:scalarpotential124810}, one ends up with such conditions, but none are usable directly to solve for two of the variables in terms of the others. The problem we are facing here is therefore a problem of symbolic regression: we are looking for analytic expressions (polynomial) that vanish once evaluated on our data points. There are a number of way one can deal with it, using already existing methods such as AI Feynman methods~\cite{Udrescu:2019mnk}, or ... \bd{litterature review}. In the spirit of trying to build a generalisable method, we develop here our own method, which is based on symbolic regression using an Annealing Importance Sampling method.

@@

		\bd{Results for naive choices of parameters (numbers of iteration and particles, probabilities, $\beta$) and motivate them (we want some exploration and then exploitation, not too long computations): for multiple runs we find multiple polynomials (good polynomial: after exploitation we convert the coefficients to integers and square roots, and recompute loss without regularisation, select with threshold). Quantify it nicely: success rates for each polynomials, and absolute number of success, failing rate. Total time needed, without cluster or fancy computers.}
	%\end{itemize}

>>>>

% ---- GRAPHIQUE ----
\usepackage{graphicx}
\usepackage{adjustbox} 
\usepackage{xcolor}
\usepackage[font=small,labelfont=bf,labelsep=space]{caption}
\usepackage{subfigure}
\captionsetup{width=0.9\textwidth,font={small,stretch=1.1}}
\addto\captionsenglish{\renewcommand{\figurename}{Fig.}}
\addto\captionsenglish{\renewcommand{\tablename}{Tab.}}
\definecolor{JoliBleu}{rgb}{0,0.55,0.55}
\definecolor{JoliVert}{rgb}{0.15,0.6,0}
\definecolor{JoliRouge}{rgb}{0.86,0.08,0}
\definecolor{JoliJaune}{rgb}{1,0.75,0}
\definecolor{JoliGris}{rgb}{0.52,0.52,0.51}
\definecolor{myblue}{RGB}{26, 77, 116}
\definecolor{myorange}{RGB}{181, 116, 30}
\definecolor{mydarkorange}{RGB}{166, 88, 0}
\definecolor{mygreen}{RGB}{21, 124, 80}
\definecolor{myblack}{RGB}{43, 65, 82}
\definecolor{myred}{rgb}{0.5, 0.0, 0.13}

@@

\title Machine learning to study marginal deformation of 2d conformal field theories

@@

The AdS/CFT correspondence \cite{Maldacena:1997re} stands as one of the most profound dualities in theoretical physics, establishing a remarkable equivalence between gravitational theories in Anti-de Sitter (AdS) spacetimes and conformal field theories (CFTs) living on their boundaries. This correspondence has revolutionized our understanding of both quantum gravity and strongly coupled field theories, providing unprecedented insights into the holographic nature of gravity.

Within this holographic framework, supergravity theories in AdS backgrounds serve as the low-energy effective descriptions of string theory compactifications, making them natural laboratories for exploring the gravitational side of the duality. Studying the CFT side is very interesting : CFT play a crucial role in statistical physics where they are known to describe phase transitions ; they are also fixed point of Renormalisation Group flow and this is this latter point that is our interest in this paper. An interesting question is indeed to know wether those CFT are isolated fixed points of RG flows, or belong to a continous family of CFT's. If this is the case, the space of deformations that takes from one CFT to another is called the conformal manifold. They are parametrized by exactely marginal operator, or in other words operators whose $\beta$ functions exactly vanish.

From this perspective, the AdS/CFT correspondence relates conformal manifolds on the boundary theory to flat directions in the scalar potential of the dual supergravity theory. Along these directions, the scalar field configurations vary continuously while the cosmological constant remains fixed. In general, supersymmetry is believed to be necessary for the existence of holographic conformal manifolds, as non-supersymmetric AdS solutions are typically considered unstable \cite{Ooguri:2016pdq,Palti:2019pca}. However, recent investigations \cite{Giambrone:2021wsm} have identified AdS$_4$ configurations that appear to evade this requirement, with no evidence of standard decay channels—neither perturbative nor non-perturbative—being present. In the context of AdS$_3$/CFT$_2$, the situation may be even richer. A well-known counterexample has existed for some time \cite{Aharony:2001dp,Dong:2014tsa}, which is well understood from both the field theory and gravitational perspectives. More generally, within the supergravity approximation, continuous deformations correspond to classically marginal operators. These operators become exactly marginal only if their marginality persists at finite values of both the rank N and the CFT coupling. Recent work \cite{Eloy:2024lwn} has demonstrated the existence of families of marginal deformations that are perturbatively stable, using the framework of Exceptional Field Theory (ExFT).

In this work, we propose a novel approach, where we study directly the scalar potential of supergravity. As already mentionned, the marginal deformation of CFT's in the large N limit, are in correspondance with the flat directions in the potential, i.e. when there is a continuous set a scalar fields such that $\nabla V = 0$. However, the explicit characterization of these flat directions presents formidable technical challenges. Supergravity scalar potentials, even in truncated models, typically involve dozens of scalar fields with intricate non-linear interactions. The resulting expressions for critical points—where all first derivatives vanish—quickly become too complex for traditional symbolic manipulation, rendering analytical approaches computationally intractable. This complexity paradox is particularly frustrating given that physically meaningful solutions often admit relatively simple parametric descriptions once discovered.

The emergence of machine learning techniques in theoretical physics opens new avenues for addressing such complex problems. Instead of solving the full symbolic system analytically from the outset, one can employ numerical methods to sample the solution space and subsequently apply symbolic regression to extract analytical patterns from the data. This hybrid approach has the potential to bypass the computational bottlenecks inherent in purely symbolic methods, while still aiming to uncover exact analytical expressions.

Similar strategies have previously been applied to identify new isolated vacua in $\mathrm{SO}(8)$ supergravity \cite{Comsa:2019rcz,Berman:2022jqn}. More broadly, there has been increasing interest in applying machine learning and numerical techniques across various domains of high-energy physics. This includes, for instance, the characterization of Calabi–Yau metrics and hypersurfaces \cite{Ashmore:2019wzb,Berman:2021mcw,Larfors:2022nep,Berglund:2022gvm,Jejjala:2020wcc,Douglas:2006rr,Larfors:2021pbb,He:2018jtw}, as well as broader efforts to explore the string theory landscape using machine learning techniques \cite{He:2017aed,Carifio:2017bov,Ruehle:2017mzq}. Additional applications include studies in CFT \cite{Chen:2020dxg}, investigations of the supergravity landscape \cite{Brady:2025zzi,Krishnan:2020sfg}, and explorations of the AdS/CFT correspondence \cite{Hashimoto:2018ftp}. More generally, machine learning has found utility in the study of string theory, geometry, and fundamental physics \cite{Ruehle:2020jrk,He:2023csq,Bao:2021auj}.

In this work, we demonstrate the viability of the aformentionned machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of three-dimensional chiral $\mathcal{N} = (1,1)$  supergravity arising from $AdS_3 \times S^3$ compactification from 6d supergravity, or from the compactification of type IIB supergravity on $AdS_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the flat direction manifold with some symbolic regression technique. There exist a large litterature on symbolic regression, using methods from genetic programming \cite{koza1994genetic,randall2022bingo,burlacu2019parsimony}, to deep machine learning \cite{petersen2019deep,kamienny2022end}, generative modelling \cite{valipour2021symbolicgpt} and diffusion models \cite{bastiani2025diffusion}. However, in the spirit of trying to build a generalisable method, we develop here our own algorithm, which is based on symbolic regression using an Annealing Importance Sampling method (AIS) \cite{neal1998annealedimportancesampling}. In this paper, we will study a 5 parameter truncation, which is itself coming from a 13 parameter consistent truncation of $\mathcal{N} = (1,1)$ 6-dimensional supergravity.

The study of this 5 dimensional parameter potentiel is separated into several parts. We first use the gradient descent to efficiently sample the underlying conformal manifold. Combined with numerical analysis, such as principal component analysis and clustering, we can identify the existence of a 3-dimensional continuous family. We then use an AIS technique combined with a sequential Monte Carlo approach (SMC) \cite{del2006sequential} in order to do symbolic regression. As we will indeed demonstrate in the paper, there exist polynomial constraints on the 5 parameters, viewed as embedding coordinates, that projects them on the 3 dimensional manifolds. We manage to identify 8 of those constraints on the data, not all independent, which once solved gives that two of the embedding coordinates can be expressed in terms of the others. This provides an explicit family of 3 dimensional parameters family

@@

As an initial simplification, we restrict our analysis to the scalar fields $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$, and $x{10}$. This subset, selected at random from the original set of thirteen scalar fields, serves to render the problem more tractable at the outset. It is important to emphasize, however, that this subset does not constitute a consistent truncation of the full theory. Consequently, the effective theory defined by these five scalars may be formally ill-defined, as higher-order interactions could, in principle, source the remaining fields.

Nevertheless, the solutions we obtain remain valid solutions of the complete theory. This is because we identify the critical points of the scalar potential by first computing $\nabla V = 0$ with all scalar fields included, and only then setting the remaining fields to zero. By performing the differentiation before the truncation, we ensure that the resulting configurations satisfy the full equations of motion and are therefore legitimate solutions of the complete theory.

To compute the gradient, we exploit the fact that we possess an analytic expression for the scalar potential \eqref{eq:scalarpotential}, which we implement in Mathematica for numerical analysis. Note however that it could have been fully done using a numerical approach, automatic differenciations methods (that we will use later to do the gradient descent for the loss function).
%Under this truncation, the scalar potential~\eqref{eq:scalarpotential} reduces to:
%\begin{equation} \label{eq:scalarpotential124810}
%	\begin{aligned}
%		V &= \frac{1}{8}\, e^{-2\,\tilde{x}_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,\tilde{x}_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
%		  & \quad\qquad\qquad - 4\,e^{3\,\tilde{x}_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
%      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
%        & \quad\qquad\qquad - 8\,e^{\tilde{x}_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
%      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
%        & \quad\qquad\qquad + 4\,e^{2\,\tilde{x}_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
%        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
%        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,\tilde{x}_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
%	\end{aligned}
%\end{equation}

Let us make a comment at this point. Since we have an explicit parametrisation of the potential with \eqref{eq:scalarpotential}, one might argue that the problem of identifying flat directions could be addressed by directly computing the gradient of the expression, setting the appropriate scalar fields to zero and subsequently simplifying the resulting equations. However, attempting to carry out this procedure using a symbolic solver such as Mathematica quickly reveals its limitations: the resulting expressions are too complex to be simplified into a manageable form, and do not yield usable relationships that express some variables in terms of others. This complexity does not, however, rule out the possibility that simpler solutions satisfying the condition of vanishing gradient may exist. In this work, we aim to identify such solutions, which we seek using numerical methods.

An additional advantage of this approach is its broader applicability. In scenarios where an explicit analytic expression for the potential and its gradient is inaccessible—yet numerical evaluation remains feasible—our method retains its validity. Hence, the numerical strategy adopted here may prove effective even in cases where an exact analytic expression for the potential is out of reach. \bd{other motivation? Example of this ? }

@@

		To do so, we perform a basic gradient descent. We begin by randomly and uniformly initializing points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range $[-1,1]$. This will be important later, when performing the symbolic regression. For this example, we choose to generate $10^5$ points. This number is motivated by the dimensionality of our problem. If we aim to populate all possible directions and want approximately $\mathcal{O}(10)$ points per direction, then we require $\mathcal{O}(10^5)$ points. Given that the true intrinsic dimensionality of the manifold is less than or equal to 5, this value serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

		We then perform gradient descent on the points using TensorFlow’s automatic differentiation. The loss function is defined as:
		\begin{equation}
			\mathcal{L} = \sum_{i = 1}^{n_{points}} ||\left.\nabla V(\vec{X}^{\,i})\,\right|_{\vec{y}=0}||^2
		\end{equation}
		where $\vec{X}^{\,i} = (x^i_1,\dots,x^i_6,\tilde{x}^i_7,\tilde{x}^i_8,\tilde{x}^i_9,x^i_{10},\dots,x^i_{12},\tilde{x}^i_{13})$ and $\vec{y}^{\,i} = (x^i_3,x^i_5,x^i_6,\tilde{x}^i_7,\tilde{x}^i_9,x^i_{11},x^i_{12},\tilde{x}^i_{11})$ denotes the data points. As already mentionned above, we use the analytic formula for $\nabla V$, but we could have perform the gradient descent by fully using automatic differentiation.

		Through trial and error, we observed that periodically reinitializing the optimizer significantly improved the convergence rate. This technique bears similarity to the concept of warm restarts introduced in \cite{loshchilov2017sgdrstochasticgradientdescent}. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence. We interpret this as follows: when the optimizer is reinitialized, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.  The evolution of the loss function, along with the learning rate schedule, is shown in Fig.~\eqref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is reinitialized. We also observe that, following the last few reinitializations, the loss exhibits a small bump immediately after the restart. We interpret this behavior as a consequence of the increased learning rate: some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence. Although this strategy introduces the risk of desampling certain regions in favor of others, they correspond to low convergence regions. As we are anyway not interested in those places, but rather real flat directions, this is not a problem for us.

@@

		The Adam optimizer was employed throughout the gradient descent procedure. It was reinitialized at iterations {250, 500, 750}, with the learning rate fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialized once more, this time with a reduced learning rate of $10^{-3}$. A final reinitialization was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 500 epochs.

		Upon completion, it was observed that all data points converged to values close to $-4$. Specifically, nearly all data points attained values within an absolute error of at least $10^{-4}$, with the exception of four points whose deviations were on the order of $10^{-3}$. Similarly, 99,739 out of 100,000 data points exhibited gradient norms smaller than $10^{-3}$. These results are summarized in Figure~\ref{fig:Analyze_V_points}.

		\begin{figure}[h!]
			\centering
			\includegraphics[scale = 0.5]{Figures/Loss_grad_des_1.png}
			\caption{Evolution of the loss function during gradient descent and corresponding learning rate schedule, noted $\alpha$ on the left y-axis, both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
			\label{fig:loss_grad_des}
		\end{figure}

		As a first visualization, we present a triangular plot of the data in Fig. \ref{triangular_plot_1_2_4_8_10}. This figure shows all possible 2D projections of the data, along with the 1D histograms of each coordinate after gradient descent.
		\begin{figure}[h!]
			\centering
			\includegraphics[scale = 0.3]{Figures/triangular_plot_1_2_4_8_10.png} %names need to be changed.
			\caption{Triangular plot showing $2d$ projections and $1d$ histograms of the data after the gradient descent}
			\label{triangular_plot_1_2_4_8_10}
		\end{figure}
		We can make a few comments on those plots. First, we observe non-trivial correlations in the data. Some of these may result from larger basins of attraction, such as in the $x_1/x_2$ plot. However, the structures seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots likely reflect genuine features of the manifold, and will show latter that this is indeed the case. We also note that all directions appear to be well populated. 
		%Furthermore, we can notice that a large majority of the points converge to the origin $\vec{0}$. This is not surprising, as this just mean that the round sphere solution has a larger basin of attraction, which sounds reasonable. We do not however attempt to shift the points along flat directions to depopulate the origin or repopulate other regions.
		%\paragraph{Local PCA: graphs of number of patches with given local dimension, and with varying size of local patch, discuss optimal patch size (we want at least few points in each direction)}
		Once the gradient descent has been completed and the flat direction sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we can first perform some exploratory analyses to better understand the data. Specifically, we aim to determine the dimensionality of the manifold and wether it consists of a single connected component or multiple disjoint components (e.g., two intersecting hyperplanes).
		To this end, we apply a local Principal Component Analysis (PCA). For each point, we identify its $k$ nearest neighbours and perform a PCA on that local neighbourhood. This procedure allows us to determine how many principal directions are needed to explain a given proportion $\epsilon$ of the data variance. In other words, it provides an estimate of the local dimensionality around each point.
		We perform this analysis for several values of $k$, namely $k \in \{5, 10, 20, 50, 100\}$, and we fix $\epsilon = 0.99$. The results are presented in Fig. \ref{local_pca}.

@@

		We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned a dimensionality of 2. This can be attributed to the fact that if the true dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality for a fraction of the points. 
		Additionally, for $k \geq 20$, we observe an increasing number of points being assigned dimensionalities of 4 or 5. This behavior can be explained by the loss of locality when the number of neighbours becomes too large: increasing $k$ results in a coarser approximation, and the algorithm may then incorporate points that are no longer truly local. This artificial enlargement of the neighbourhood can cause the estimated local dimensionality to rise.
		We thus conclude that the manifold under investigation has an intrinsic dimension of 3.

@@

		Now we have sampled the manifold and extracted some basic information about it (namely its dimension and the fact that it is made of one block), we would like to see if we can extract some analytic formula to characterise it. What we have at the moment, are points on a 3-dimensional manifold, which are embedded in a 5-dimensional space. Therefore we can conclude that, in order to characterise the manifold, we need two independent constraints on the embedding coordinates $\vec{x} = (x_1,x_2,x_4,\tilde{x}_8,x_{10})$. If we have a look at the form of the potential, we can observe that if we use $x_{8} = e^{\tilde{x}_8}$, that, up to some potential global factor of the dilaton of the form $e^{\tilde{x}_8}$, this potential is actually a polynomial of the embedding coordinates. Therefore, the components of $|\nabla V|$ are also polynomials in those variables. We conclude that the constraints on the embedding coordinates we are looking for are polynomial constraints of the form $p(\vec{x}) = 0$, and that there should be at least two of those. Note that we are now taking $\vec{x} = (x_1,x_2,x_4,x_8,x_{10})$. Of course if one takes directly the gradient of \eqref{eq:scalarpotential}, one ends up with such conditions, but none are usable directly to solve for two of the variables in terms of the others. The problem we are facing here is therefore a problem of symbolic regression: we are looking for analytic expressions (polynomial) that vanish once evaluated on our data points.

@@

%		\bd{Results for naive choices of parameters (numbers of iteration and particles, probabilities, $\beta$) and motivate them (we want some exploration and then exploitation, not too long computations): for multiple runs we find multiple polynomials (good polynomial: after exploitation we convert the coefficients to integers and square roots, and recompute loss without regularisation, select with threshold). Quantify it nicely: success rates for each polynomials, and absolute number of success, failing rate. Total time needed, without cluster or fancy computers.}
	%\end{itemize}
